---
title: "Volcano Plots"
#output: html_notebook
---

This worksheet will produced Volcanoplots based on the differential gene results on the scRNAseq dataset in the 2023_08_29_tonsil_scRNA project.

#Load libraries

```{r}
library(EnhancedVolcano)
```

# Read in data

```{r}
m1.mac <- read.csv(file = "Analysis_2023_08_29/results/m1_mac_Sig_Genes.csv")
m1.mac.deseq <- read.csv(file = "Analysis_2023_08_29/results/m1_mac_DESeq2_Genes.csv")
ilc1 <- read.csv(file = "Analysis_2023_08_29/results/ilc1_Sig_Genes.csv")
ilc3 <- read.csv(file = "Analysis_2023_08_29/results/ilc3_Sig_Genes.csv")
nk.16neg <- read.csv(file = "Analysis_2023_08_29/results/nk_cd16_neg_Sig_Genes.csv")
nk.16pos <- read.csv(file = "Analysis_2023_08_29/results/nk_cd16_pos_Sig_Genes.csv")
pdc <- read.csv(file = "Analysis_2023_08_29/results/pdc_Sig_Genes.csv")
dc1 <- read.csv(file = "Analysis_2023_08_29/results/dc1_Sig_Genes.csv")
dc2 <- read.csv(file = "Analysis_2023_08_29/results/dc2_Sig_Genes.csv")
dc5 <- read.csv(file = "Analysis_2023_08_29/results/dc5_Sig_Genes.csv")
mmp12.mac <- read.csv(file = "Analysis_2023_08_29/results/mmp12_mac_Sig_Genes.csv")
ptgd5.mac <- read.csv(file = "Analysis_2023_08_29/results/ptgd5_mac_Sig_Genes.csv")
adc1 <- read.csv(file = "Analysis_2023_08_29/results/aDC1_Sig_Genes.csv")
neutro <- read.csv(file = "Analysis_2023_08_29/results/neutro_Sig_Genes.csv")
mono <- read.csv(file = "Analysis_2023_08_29/results/mono_Sig_Genes.csv")

```


# Volcano Plots

## M1 Macrophages

```{r}
EnhancedVolcano(m1.mac,
                lab = m1.mac$symbols,
                x = "avg_log2FC",
                y = "p_val_adj",
                title = "Downs Syndrome vs Healthy Control",
                subtitle = "M1 Macrophage population")
ggsave(filename = "Analysis_2023_09_06/m1.mac.volcano.png",
       width = 11,
       height = 8,
       units = "in")
```

```{r}
EnhancedVolcano(m1.mac.deseq,
                lab = m1.mac.deseq$symbols,
                x = "avg_log2FC",
                y = "p_val_adj",
                title = "Downs Syndrome vs Healthy Control",
                subtitle = "DESeq2: M1 Macrophage population")
ggsave(filename = "Analysis_2023_09_06/m1.mac.deseq.volcano.png",
       width = 11,
       height = 8,
       units = "in")
```

## ILC1

```{r}
EnhancedVolcano(ilc1,
                lab = ilc1$symbols,
                x = "avg_log2FC",
                y = "p_val_adj",
                title = "Downs Syndrome vs Healthy Control",
                subtitle = "ILC1 population")
ggsave(filename = "Analysis_2023_09_06/ilc1.volcano.png",
       width = 11,
       height = 8,
       units = "in")
```

## ILC3

```{r}
EnhancedVolcano(ilc3,
                lab = ilc3$symbols,
                x = "avg_log2FC",
                y = "p_val_adj",
                title = "Downs Syndrome vs Healthy Control",
                subtitle = "ILC3 population")
ggsave(filename = "Analysis_2023_09_06/ilc3.volcano.png",
       width = 11,
       height = 8,
       units = "in")
```

## CD16- NK cells

```{r}
EnhancedVolcano(nk.16neg,
                lab = nk.16neg$symbols,
                x = "avg_log2FC",
                y = "p_val_adj",
                title = "Downs Syndrome vs Healthy Control",
                subtitle = "CD16- NK cells")
ggsave(filename = "Analysis_2023_09_06/nk.16neg.volcano.png",
       width = 11,
       height = 8,
       units = "in")
```

## CD16+ NK cells

```{r}
EnhancedVolcano(nk.16pos,
                lab = nk.16pos$symbols,
                x = "avg_log2FC",
                y = "p_val_adj",
                title = "Downs Syndrome vs Healthy Control",
                subtitle = "CD16+ NK cells")
ggsave(filename = "Analysis_2023_09_06/nk.16pos.volcano.png",
       width = 11,
       height = 8,
       units = "in")
```

## pDC cells

```{r}
EnhancedVolcano(pdc,
                lab = pdc$symbols,
                x = "avg_log2FC",
                y = "p_val_adj",
                title = "Downs Syndrome vs Healthy Control",
                subtitle = "pDC cells")
ggsave(filename = "Analysis_2023_09_06/pdc.volcano.png",
       width = 11,
       height = 8,
       units = "in")
```

## DC1 Mature cells

```{r}
EnhancedVolcano(dc1,
                lab = dc1$symbols,
                x = "avg_log2FC",
                y = "p_val_adj",
                title = "Downs Syndrome vs Healthy Control",
                subtitle = "DC1 Mature cells")
ggsave(filename = "Analysis_2023_09_06/dc1.mature.volcano.png",
       width = 11,
       height = 8,
       units = "in")
```

## DC2 cells

```{r}
EnhancedVolcano(dc2,
                lab = dc2$symbols,
                x = "avg_log2FC",
                y = "p_val_adj",
                title = "Downs Syndrome vs Healthy Control",
                subtitle = "DC2 cells")
ggsave(filename = "Analysis_2023_09_06/dc2.volcano.png",
       width = 11,
       height = 8,
       units = "in")
```

## DC5 cells

```{r}
EnhancedVolcano(dc5,
                lab = dc5$symbols,
                x = "avg_log2FC",
                y = "p_val_adj",
                title = "Downs Syndrome vs Healthy Control",
                subtitle = "DC5 cells")
ggsave(filename = "Analysis_2023_09_06/dc5.volcano.png",
       width = 11,
       height = 8,
       units = "in")
```

## aDC1 cells

```{r}
EnhancedVolcano(adc1,
                lab = adc1$symbols,
                x = "avg_log2FC",
                y = "p_val_adj",
                title = "Downs Syndrome vs Healthy Control",
                subtitle = "DC5 cells")
ggsave(filename = "Analysis_2023_09_06/adc1.volcano.png",
       width = 11,
       height = 8,
       units = "in")
```

## IL7R MMP12 macrophages

```{r}
EnhancedVolcano(mmp12.mac,
                lab = mmp12.mac$symbols,
                x = "avg_log2FC",
                y = "p_val_adj",
                title = "Downs Syndrome vs Healthy Control",
                subtitle = "IL7R MMP12 macrophages")
ggsave(filename = "Analysis_2023_09_06/mmp12.mac.volcano.png",
       width = 11,
       height = 8,
       units = "in")
```

## SELENOP FUCA1 PTGDS macrophages

```{r}
EnhancedVolcano(ptgd5.mac,
                lab = ptgd5.mac$symbols,
                x = "avg_log2FC",
                y = "p_val_adj",
                title = "Downs Syndrome vs Healthy Control",
                subtitle = "SELENOP FUCA1 PTGDS macrophages")
ggsave(filename = "Analysis_2023_09_06/ptgd5.mac.volcano.png",
       width = 11,
       height = 8,
       units = "in")
```

## Monocytes

```{r}
EnhancedVolcano(mono,
                lab = mono$symbols,
                x = "avg_log2FC",
                y = "p_val_adj",
                title = "Downs Syndrome vs Healthy Control",
                subtitle = "Monocytes")
ggsave(filename = "Analysis_2023_09_06/mono.volcano.png",
       width = 11,
       height = 8,
       units = "in")
```

## Neutrophil Granulocytes

```{r}
EnhancedVolcano(neutro,
                lab = neutro$symbols,
                x = "avg_log2FC",
                y = "p_val_adj",
                title = "Downs Syndrome vs Healthy Control",
                subtitle = "Neutrophils Granulocytes")
ggsave(filename = "Analysis_2023_09_06/neutro.volcano.png",
       width = 11,
       height = 8,
       units = "in")
```























