---
title: "2024_05_22_Elena_plot_requests"
format: html
editor: source
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r libraries}
suppressPackageStartupMessages({
library(Seurat)
library(tidyverse)
library(qs)
})
```

```{r}
results.dir <- "Analysis_2024_05_22/"
```


# Read in data: Post Anchor SCT Object

```{r read in post anchor seurat}
SCT_data_object = qread("data/2023_08_29_SCT_data_object_Post_Anchored.qs", nthreads = 10)
```


```{r}
ds_samples <- c("ds_tn_67", "ds_tn_6865")
#hc_samples <- c("hc_65", "hc_69")

SCT_data_object@meta.data <- SCT_data_object@meta.data |> 
  dplyr::mutate(group = ifelse(orig.ident %in% ds_samples, "DS", "NDS"),
                .after = orig.ident)
```

# Subset data

```{r}
# Switch Idents to group
# Idents(SCT_data_object) <- "group"

# ds_object <- subset(SCT_data_object, idents = c("ds"))
# hc_object <- subset(SCT_data_object, idents = c("hc"))

# return Idents to predicted.id
# Idents(ds_object) <- "predicted.id"
# Idents(hc_object) <- "predicted.id"
```

# Plot Data
## Original plots

```{r}
VlnPlot(SCT_data_object,
        features = "IL1B",
        group.by = "predicted.id",
        sort = "increasing") + 
  theme(legend.position = 'none',
        axis.text = element_text(size = 8))
```

```{r}
VlnPlot(hc_object,
        features = "IL1B",
        group.by = "predicted.id",
        sort = "increasing") + 
  theme(legend.position = 'none',
        axis.text = element_text(size = 8)) +
        labs(title = "IL1B, NDS samples")
```

# How to fix "Low Resolution" problem

## Increase DPI

```{r}
VlnPlot(hc_object,
        features = "IL1B",
        group.by = "predicted.id",
        sort = "increasing") + 
  theme(legend.position = 'none',
        axis.text = element_text(size = 8)) +
        labs(title = "IL1B, NDS samples")

ggsave(filename = paste0(results.dir,"IL1B.NDS.300.png"),
       width = 11,
       height = 8,
       units = "in",
       dpi = 300)

ggsave(filename = paste0(results.dir,"IL1B.NDS.600.png"),
       width = 11,
       height = 8,
       units = "in",
       dpi = 600)
```

## Cut down on number of predicted.ids

```{r}
hc_object@meta.data |> 
  group_by(predicted.id) |> 
  summarise(count = n())|> 
  pull(predicted.id)
```

```{r}
shorter.vector <- c("M1 Macrophages", "Neutrophil Granulocytes", "DC4", "C1Q HLA macrophages", "DC2", "Monocytes")
```

Make sure that the Idents() of your single cell object is on whatever metadata column you are creating a subset vector from!!!

```{r}
VlnPlot(hc_object,
        features = "IL1B",
        group.by = "predicted.id",
        sort = "increasing",
        idents = shorter.vector) + 
  theme(legend.position = 'none',
        axis.text = element_text(size = 8)) +
        labs(title = "IL1B, NDS samples")

# ggsave(filename = paste0(results.dir,"IL1B.shorter.600.png"),
#        width = 11,
#        height = 8,
#        units = "in",
#        dpi = 600)
```

## Split plot

```{r}
VlnPlot(SCT_data_object,
        features = "IL1B",
        group.by = "predicted.id",
        sort = "increasing",
        split.by = "group") + 
  theme(axis.text = element_text(size = 8))

# ggsave(filename = paste0(results.dir,"IL1B.split.600.png"),
#        width = 11,
#        height = 8,
#        units = "in",
#        dpi = 600)
```

## 2024_05_24 Update

Showed Elena the previous plots and we are moving forward with a grouped plot consisting of only the myeloid subsets.

```{r}
rm(ds_object, hc_object, il1b.sub)
```

She wants the following groups: M1, Neut-Gran, C1Q MAC, Monocytes, IL7R Macs, ITGAX ZEBS Macs.

```{r}
SCT_data_object@meta.data |> 
  group_by(predicted.id) |> 
  summarise(count = n()) |> 
  pull(predicted.id)
```

```{r}
myeloid.vector <- c("M1 Macrophages", "Neutrophil Granulocytes", "C1Q HLA macrophages", "Monocytes", "IL7R MMP12 macrophages", "ITGAX ZEB2 macrophages" )

# Earlier changed SCT_data_object Idents() to group, need to change it back to plot properly.
Idents(SCT_data_object) <- "predicted.id"
```

```{r}
VlnPlot(SCT_data_object,
        features = "IL1B",
        group.by = "predicted.id",
        sort = "increasing",
        split.by = "group",
        idents = myeloid.vector) + 
  theme(axis.text = element_text(size = 8))
```

This is the plot we want. Need to clean it up a bit more and change the order of display if we can.


```{r}
class(SCT_data_object@meta.data$group)
```

```{r}
# change order of DS/NDS on plots
SCT_data_object@meta.data$group <- as.factor(SCT_data_object@meta.data$group)

levels(SCT_data_object@meta.data$group)

```

```{r}
SCT_data_object@meta.data$group <- relevel(SCT_data_object@meta.data$group, ref = "NDS")
```

```{r}
VlnPlot(SCT_data_object,
        features = "IL1B",
        group.by = "predicted.id",
        sort = "increasing",
        split.by = "group",
        idents = myeloid.vector) + 
  theme(axis.text = element_text(size = 8))
```

Okay we have swapped them, but the colors swapped too. 

```{r}
VlnPlot(SCT_data_object,
        features = "IL1B",
        group.by = "predicted.id",
        sort = "increasing",
        split.by = "group",
        idents = myeloid.vector,
        cols = c("blue", "red")) + 
  theme(axis.text = element_text(size = 7))
```

```{r}
ggsave(filename = paste0(results.dir,"IL1B.grouped.900dpi.png"),
       width = 11,
       height = 8,
       units = "in",
       dpi = 900)
```

### CD127/IL7R gene in other subsets

Elena also asked for "cd127 in NKS,ILCS and DCS between DS and NDS".

Let's pull the group names again for easy building of our idents vector

```{r}
SCT_data_object@meta.data |> 
  group_by(predicted.id) |> 
  summarise(count = n()) |> 
  pull(predicted.id)
```

```{r}
cd127.vector <- c("CD16-CD56- NK", "DC2", "DC5", "ILC1", "NKp44+ ILC3", "CD16+CD56- NK", "DC1 mature", "DC4", "NKp44- ILC3", "PDC", "aDC1")
```


```{r}
VlnPlot(SCT_data_object,
        features = "IL7R",
        group.by = "predicted.id",
        sort = "increasing",
        split.by = "group",
        idents = cd127.vector,
        cols = c("blue", "red")) + 
  theme(axis.text = element_text(size = 7))
```
```{r}
ggsave(filename = paste0(results.dir,"CD127.grouped.900dpi.png"),
       width = 11,
       height = 8,
       units = "in",
       dpi = 900)
```





