---
title: "2024_06_27_EIF2_gene_set"
format: html
editor: source
---

Elena asked about some EIF signaling data in some subsets of the single cell data, so I will investigate that gene set in this notebook.

# Directores

```{r dir folders}
folder.name <- "Analysis_2025_06_24"

if(!exists(folder.name)){
  dir.create(folder.name)
}

```

# Libraries

```{r libraries}
suppressPackageStartupMessages({
library(Seurat)
library(tidyverse)
library(org.Hs.eg.db)
library(clusterProfiler)
library(ReactomePA)
library(gprofiler2)
library(DESeq2)
library(msigdbr)
library(qs)
})
```

# Import Data

The gene sets I sent are in Analysis_2023_08_29. Some additional samples have been added since then, in 2023_10_03 and two others I have not yet added to the object. So I may need to revisit this notebook after doing that. For now we will continue with the original data.

```{r}
monocyte.data <- read.csv(file = "Analysis_2023_08_29/results/mono_Sig_Genes.csv") |> 
  dplyr::filter(p_val_adj < 0.05)
ilc1.data <- read.csv(file = "Analysis_2023_08_29/results/ilc1_Sig_Genes.csv") |> 
  dplyr::filter(p_val_adj < 0.05)
ilc3.data <- read.csv(file = "Analysis_2023_08_29/results/ilc3_Sig_Genes.csv") |> 
  dplyr::filter(p_val_adj < 0.05)
cd16pos.nk.data <- read.csv(file = "Analysis_2023_08_29/results/nk_cd16_pos_Sig_Genes.csv") |> 
  dplyr::filter(p_val_adj < 0.05)
cd16neg.nk.data <- read.csv(file = "Analysis_2023_08_29/results/nk_cd16_neg_Sig_Genes.csv") |> 
  dplyr::filter(p_val_adj < 0.05)
```

# EIF2 signaling pathway

Now we need to find the gene set for EIF2 signaling

```{r}
mono.go <- groupGO(gene = monocyte.data$symbols,
                   keyType = "SYMBOL",
                   OrgDb = org.Hs.eg.db,
                   ont = "BP",
                   level = 2,
                   readable = T)
head(mono.go)
```


```{r}
mono.gse <- gseGO(geneList = monocyte.data$symbols,
                  keyType = "SYMBOL",
                  OrgDb = org.Hs.eg.db,
                  ont = "BP",
                  verbose = FALSE)

head(mono.gse)
```















